### 2015-11-10 ResearchStatement/main.tex
First full draft

\noindent
For many problems in many disciplines, it is easier to determine the quality of a solution
than to find the best possible solution. For example, this is true of NP-Hard problems,
finite element analysis, chemical design, and physics simulations.
While coming from very different fields, these problems share
three critical characteristics:
\begin{itemize}
\item There is a defined set of possible solutions.
\item We can measure the quality of a proposed solution.
\item There is no tractable method for finding the best solution.
\end{itemize}
I design algorithms for optimizing problems with these characteristics. Instead of
attempting to create specialized heuristics for every possible problem,
my goal is to discover and utilize general principles which can
then be applied to a large set of problems without further expert knowledge.

@@

~\cite{goldman:2015:GBO}

~\cite{goldman:2016:hyperplane}

~\cite{whitley:2016:mkl}
TODO Exact polynomial time

TODO SAT compression

@@

TODO Statistical Physics somewhere

@@

TODO Neuro-evolution

\section{Future Work}
TODO Chemistry (Proteins / Batteries)

TODO Multi-objectiveness

TODO Generalized search space compression

>>>>

\noindent
For many problems in many disciplines, it is easier to determine the quality of a solution
than to find the best possible solution. For example, this is true of NP-Hard problems,
finite element analysis, chemical design, and physics simulations.
While coming from very different fields, these problems share
three critical characteristics:
\begin{itemize}
\item There is a defined set of possible solutions.
\item We can measure the quality of a proposed solution.
\item There is no tractable method for finding the best solution.
\end{itemize}
I design algorithms for optimizing problems with these characteristics. Instead of
attempting to create specialized heuristics for every problem,
my goal is to discover and utilize general principles which can
then be applied to a large set of problems without further expert knowledge.

@@

By making slightly more information about the problem available to optimization, I was
able to develop an even more efficient version of P3~\cite{goldman:2015:GBO}. Consider
that many problems can be expressed as the sum of functions which use only a subset
of problem variables. For example, Boolean Satisfiability is often expressed as the sum
of clauses. The Maximum Cut of a graph is equal to the sum of checking each edge.
Using this additional information, I was able to reduce P3's runtime by two
orders of magnitude. I was also able to construct an efficient method
for finding all local optima in an arbitrary problem~\cite{goldman:2016:hyperplane},
and with a collaborator create a generic method which solves many common
problems in the optimization literature in $O(N)$ time~\cite{whitley:2016:mkl}.

@@

%TODO Statistical Physics somewhere

@@

A challenging task in machine learning is the training of large scale neural networks.
I have recently been in collaboration with an evolutionary biologist
who studies natural learning to develop better methods for training artificial brains.

\section{Future Work}
In recently submitted work~\cite{whitley:2016:mkl}, we determined that many real-world
instances of common NP-Hard problems contained a surprising amount of regularity.
I believe this regularity can be exploited to significantly reduce the space
which needs to be searched for high quality solutions. To that end, I plan to develop
a problem independent technique for search space compression.

A current limitation of P3 is the focus on single objective optimization. For many
problems, there are many competing objectives which should be optimized. For example,
strength, weight, and cost in an engineering task. Optimization gains significant
benefits from treating this problem as having three separate goals instead of attempting
to combine them into a single objective. I believe that integrating P3 with the concepts
of multi-objective optimization will result in a powerful tool for real-world optimization.

Finally, and most critically, I plan to continue exploring interdisciplinary collaborations.
As a tool for problem solving, optimization algorithms are only as useful as the problems they
can solve. Some specific targets for future collaborations include better battery design,
structural engineering tasks, and land use management.
